Tesseroids: command-line tools for gravity field forward modeling in spherical coordinates

### 2014-04-01 manuscript.tex
Sect about shell comparison comes before software
Finished the theory section before and started on the shell comparison
section.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
    \caption{
        Adaptive discretization
        of the tesseroid shown in (a)
        for a computation point P
        using distance/size ratios of
        (b) 0.5, (c) 1, and (d) 2.
        $d$ is the distance between the tesseroid and P.
    }
    \label{fig:split}
\end{figure}

@@

The potential of a spherical half-shell is

\begin{equation}
    V(r) = 2\pi G \rho \left[ \dfrac{l^3 + {r'}^3}{3r} - 0.5 {r'}^2 \right]
           \Biggr \rvert_{r'=r_1}^{r'=r_2}
    \label{eq:halfshell-pot}
\end{equation}

>>>>

We use a modified version
of the recursive scheme of \citet{Li2011}.
We fix the number of nodes in the GLQ
$N^\lambda=N^\phi=N^r=2$.
In this case, the maximum distance between nodes
is approximately the largest dimension of the tesseroid ($L$)
\citep{Wild-Pfeiffer2008}.
For a computation point $P$
at a distance $d$ from
the center of the top face of the tesseroid
(Figure~\ref{fig:ratio}a),
the tesseroid is divided in eight parts if

\begin{equation}
    \frac{d}{L} < R,
    \label{eq:ratio}
\end{equation}

\noindent
where $R$ is the minimum distance/size ratio desired.
In case the tesseroid is divided,
the inequality in equation \ref{eq:ratio}
is checked for each of the eight parts.
This process is repeated recursively
until no divisions are required,
ensuring that equation \ref{eq:ratio}
holds for all tesseroids created.
The gravitational effect
of the original tesseroid
is then calculated as
the sum of the effect
of the smaller tesseroids.

The rule-of-thumb of \citet{Ku1977}
used by \citet{Li2011}
is equivalent to setting $R=1$.
Figure~\ref{fig:ratio}
illustrates the subdivision of a tesseroid
(Figure~\ref{fig:ratio}a)
for computation point $P$
using increasing values of $R$.
Note that increasing $R$
results is a finer division of the tesseroid
around the computation point
and thus a more accurate GLQ integration.
However, higher values of $R$
also result in
larger computation times.
For example,
the tesseroid in Figure~\ref{fig:ratio}a
is replaced by
50 tesseroids for $R=0.5$ (Figure~\ref{fig:ratio}b),
344 tesseroids for $R=1$ (Figure~\ref{fig:ratio}c),
and 2472 tesseroids for  $R=2$ (Figure~\ref{fig:ratio}d).

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/tesseroid-split}
    \caption{
        Adaptive discretization
        of the tesseroid shown in (a)
        for a computation point P
        using $R$ equal to
        (b) 0.5, (c) 1, and (d) 2.
        $d$ is the distance between
        the center of the top face of the tesseroid
        and $P$.
        $L$ is the largest dimension of the tesseroid.
    }
    \label{fig:ratio}
\end{figure}

Clearly,
there is need to determine
an optimal value of $R$
that yields the desired accuracy
in the GLQ integration.
In the following section,
we compare the computed gravitational effects
of a tesseroid model
with that of a spherical half-shell
(Figure~\ref{fig:shell}),
for which exist analytical expressions.

@@

The gravitational potential
of a homogeneous spherical half-shell
at a computation point P
located at a height $h$
from the center of the shells top face
(Figure~\ref{fig:shell})
is

\begin{equation}
    V(r) = 2\pi G \rho \left[ \dfrac{l^3 + {r'}^3}{3r} - 0.5 {r'}^2 \right]
           \Biggr \rvert_{r'=r_1}^{r'=r_2},
    \label{eq:halfshell-pot}
\end{equation}
